---
title: "Reproducible Research Project 1"
author: "Robert Ramers"
date: "March 11, 2016"
output: html_document
---

# Getting and processing data
Load library for the project:  

```{r}
library(ggplot2) #to do plots  
```

## 1. Load the data  

```{r}
data.1 <- read.csv("activity.csv")
```  


## Clean data, remove NA's
```{r}
dataf <-na.omit(data.1)
```

## Question 1: Mean number of steps taken per day
First, we will make a histogrm of the total number of steps taken each day

```{r}
dataf.steps <- aggregate(steps ~ date,dataf,sum)
hist(dataf.steps$steps, col='darkgreen',main = "Histogram of steps per day",xlab="Total Steps  per day")
```
Now we will find the mean and the median of steps per day
```{r}
rmean <- mean(dataf.steps$steps)
rmedian <- median(dataf.steps$steps)
rmean
rmedian

```
##Question 2: Calculate the daily activity pattern
```{r, echo=FALSE}
dataf.averages <- aggregate(x=list(steps=dataf$steps),by=list(interval=dataf$interval),FUN=mean)
ggplot(data=dataf.averages, aes(x=interval, y= steps)) +geom_line()+xlab("Five minute intervals")+ylab("Average steps")
```


To determine the five minute interal with the max steps across all days:
```{r}
maxinterval <- dataf.averages[which.max(dataf.averages$steps),1]
maxinterval
```
## Imputing missing values
We now need to impute data for the missing values (NA) by inserting the average for each interval.
```{r}
incomplete <- sum(!complete.cases(data.1))
incomplete
imputed_values <- transform(data.1, steps=ifelse(is.na(data.1$steps),dataf.averages$steps[match(dataf$interval, dataf.averages$interval)],dataf$steps))
```
Recount the total steps and make histogram:

```{r}
dataf.steps1 <-aggregate(steps ~date,imputed_values,sum)
hist(dataf.steps1$steps,main = paste("Total steps per day"),col="darkgreen",xlab = "Number of Steps")
```

Create a histogram that shows the difference:

```{r}
hist(dataf.steps1$steps,main = paste("Total steps per day"),col="darkgreen",xlab = "Number of Steps")
hist(dataf.steps$steps,main = paste("Total steps per day"),col="red",xlab="Number of steps", add=T)
legend("topright",c("Imputed","Non imputed"),col=c("darkgreen","red"),lwd=10)
```


Compute mean and median for imputed data:  
```{r}
mean.i <- mean(dataf.steps1$steps)
median.i <- median(dataf.steps1$steps)
mean.i
median.i
```
Calculate difference - imputed vs. non imputed data:  
```{r}
meandiff <-mean.i - rmean
mediff <- median.i -  rmedian
meandiff
mediff
```
Calculate the total difference:  
```{r}
difference <- sum(dataf.steps1$steps) - sum(dataf.steps$steps)
difference
```

#Differences in activity patterns between weekends and weekdays

```{r}
weekdays <- c("Monday","Tuesday", "Wednesday","Thursday","Friday")
imputed_values$dow=as.factor(ifelse(is.element(weekdays(as.Date(imputed_values$date)),weekdays),"weekday","weekend"))
dataf.steps1 <- aggregate(steps~ interval+dow,imputed_values,mean)
library(lattice)
xyplot(dataf.steps1$steps ~ dataf.steps1$interval|dataf.steps1$dow,main="Average steps per day by interval",xlab="Interval",ylab="Steps",layout=c(1,2),type="1")
```

